### 2023-02-21 figures/figures.Rmd
repo infra

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
- \pagenumbering{gobble}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = "H")
```

\pagebreak

```{r}
source("../scripts/99_utils.R")
```

```{r map}
print(2)
```