### 2023-02-21 figures/figures.Rmd
data loading

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
- \pagenumbering{gobble}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = "H")
```

\pagebreak

```{r}
if(dir.exists("../scripts")){
    source("../scripts/99_utils.R")
}else{
    source("scripts/99_utils.R")
}

dt <- read.csv(
    "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt",
    header = TRUE,
    colClasses = c("TIMESTAMP_START" = "character", "TIMESTAMP_END" = "character"), 
    na.strings = "-9999") %>%
    clean_names()

head(dt)
```

```{r map}
print(2)
```