### 2023-03-06 figures/figures.Rmd
print distances

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{colortbl}
- \usepackage{float}
- \usepackage{caption}
- \pagenumbering{gobble}
---

\arrayrulecolor{white}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", dev = c("pdf"))
knitr::opts_chunk$set(fig.pos = "H")
```

\pagebreak

```{r, echo=FALSE}
if(dir.exists("../scripts")){
    shim <- "../"
    source("../scripts/99_utils.R")
}else{
    shim <- ""
    source("scripts/99_utils.R")
}
```

```{r}
dt <- read.csv(
    paste0(shim, "../../Data/Euroflux/BE-Lon/L2-L4_2004-2012/EFDC_L2_Flx_BELon_2008_v017_30m.txt"),
    header = TRUE,
    colClasses = c("TIMESTAMP_START" = "character", "TIMESTAMP_END" = "character"),
    na.strings = "-9999") %>%
    clean_names()

head(dt)
```

```{r map}

coords_ire <- sf::st_as_sf(
    data.frame(lat = 50.45055, lon = 4.5350415),
    coords = c("lon", "lat"), crs = 4326)

dt_sites <- read.csv(
    system.file("extdata", "Site_metadata.csv", package = "FluxnetLSM")) %>%
    clean_names() %>%
    dplyr::mutate_all(na_if, "") %>%
    st_as_sf(coords = c("site_longitude", "site_latitude"), crs = 4326)

dt_sites$dist <- unlist(list(unlist(st_distance(coords_ire, dt_sites)))) / 1609.34
dt_sites_close <- dt_sites[dt_sites$dist < 100,]
head(dplyr::arrange(dt_sites_close, dist)$dist)

rmarkdown::render_delayed({
    mapview(dt_sites_close) + mapview(coords_ire, color = "red")
    })
```